Get Your Own Just-In-Time Specializing Compiler For Free

### 2007-05-29 talk/dls2007/paper.tex
More details.

\title{Generating Just-In-Time Specializing Compilers}

@@

\begin{abstract}
PyPy's translation tool-chain -- from the interpreter written in RPython
to generated VMs for low-level platforms -- is now able to extend those
VMs with an automatically generated dynamic compiler, derived from the
interpreter. This is achieved by a pragmatic application of partial
evaluation techniques guided by a few hints added to the source of the
interpreter. Crucial for the effectiveness of dynamic compilation is
the use of run-time information to improve compilation results: in
our approach, a novel powerful primitive called "promotion" that "promotes"
run-time values to compile-time is used to that effect.  In this report,
we describe it along with other novel techniques that allow the approach
to scale to something as large as PyPy's Python interpreter.
\end{abstract}

@@

.. comment

    python 2.4:                     0.82
    pypy-c-42412-allworking-tproxy: 1.90392203331
    pypy-c-41802-faassen:           1.60267777443
    pypy-c-41992-jit, turned off:   1.76742062569
    pypy-c-41992-jit:               0.00912307977676   <===
    psyco 1.5.2:                    0.00621596813202
    gcc -O0:                        0.0061992
    gcc -O2:                        0.0021689

@@

>>>>

\title{Get Your Own Just-In-Time Specializing Compiler For Free}

@@

\begin{abstract}
PyPy's translation tool-chain -- from the interpreter written in RPython
to generated VMs for low-level platforms -- is now able to extend those
VMs with an automatically generated dynamic compiler, derived from the
interpreter. This is achieved by a pragmatic application of partial
evaluation techniques guided by a few hints added to the source of the
interpreter. Crucial for the effectiveness of dynamic compilation is
the use of run-time information to improve compilation results: in
our approach, a novel powerful primitive called "promotion" that "promotes"
run-time values to compile-time is used to that effect.  In this report,
we describe it along with other novel techniques that allow the approach
to scale to something as large as PyPy's Python interpreter.
\footnote{This research was partially supported by the EU funded
 project: IST 004779 PyPy (PyPy: Implementing Python in Python).}
\end{abstract}

@@

@@

\bigskip

\bibliographystyle{abbrv}
\bibliography{paper}
